Clearance Project - 1 - Data Generation

### 2025-03-19 data-raw/ClearanceProject-1-DataRmd.Rmd
Updated DESCRIPTION and data.

@@

@@

# ctxR download of all chemicals in 3comp2 model
# Chemical info and properties (averaging the latter by chemical and experimental/predicted)
chem_info <- get_chemical_details_batch(DTXSID = all_chems,
                                               rate_limit = 0L) %>%
  filter(!is.na(smiles))
# 9086 total unique chemicals with SMILES data

@@

functional_uses_all <- readr::read_csv(paste0(
  Sys.getenv("FIG_DIR"), "RestrictiveClearance_Analysis/",
  "ChemExpo_bulk_functional_uses_20250304.csv"
))

@@

@@

>>>>

This code chunk creates and `.RData` file that contains character vectors
of DTXSIDs for:

1. All chemicals with human parameterization available for a 3-compartment model w/ inhalation.  
2. All chemicals with rat parameterization available for a 3-compartment model w/ inhalation.  
3. All chemicals with human and rat plasma concentration over time data.

@@

Below read in some external data which is publicly available already.

- `red_assay_res` is the named object that contains the data from the RED assay
including the dissociation constant $K_d$ for plasma protein binding.  
- `chem_info` is a combined data.frame from calls to `get_chemical_details_batch`
and `get_chem_info_batch`. This uses the `ctxR` API to fetch chemical information
such as physio-chemical properties and SMILES from the CompTox dashboard.  
- `cyp_data` is data from Sipes 2017 predicting the CLint fraction for each CYP enzyme.
- `functional_uses_all` is a list of functional use categories for each chemical
from the Chemical and Products Database.

@@

# ctxR download of all chemicals in 3comp2 model
# Chemical info and properties (averaging the latter by chemical and experimental/predicted)
chem_info <- get_chemical_details_batch(DTXSID = human_3compchems,
                                               rate_limit = 0L) %>%
  filter(!is.na(smiles))
# 9086 total unique chemicals with SMILES data

@@

# Sipes 2017 data
cyp_data <- readr::read_csv(paste0(
  Sys.getenv("FIG_DIR"),
  "RestrictiveClearance_Analysis/",
  "CLint_CYP_Sipes2017.csv"
)) %>%
  left_join(
    chem.physical_and_invitro.data %>%
      select(CAS = CAS, Chemical = DTXSID) %>%
      distinct()
  ) %>%
  select(Chemical, CAS, starts_with("MET_")) %>%
  filter(!is.na(Chemical))
names(cyp_data) <- c("Chemical", "CAS", "CYP1A2",
                     "CYP2C9", "CYP2C19", "CYP2D6", "CYP3A4")
# Functional use categories from CPdat
functional_uses_all <- readr::read_csv(paste0(
  Sys.getenv("FIG_DIR"), "RestrictiveClearance_Analysis/",
  "ChemExpo_bulk_functional_uses_20250304.csv"
))

save(cyp_data,
     file = "data-raw/CYP_Sipes2017.Rdata")

@@

Next we fit the plasma data from CvTdb, which we already have in the `invivoPKfit`
package using `cvt` object. We also get a few downstream statistics and save them
in an `.RData` file.

@@

The next two chunks take a while. These will parameterize the `3compartment2`
`httk` model with restrictive and nonrestrictive clearance.
It also creates data.frames for evaluating predictions, steady-state concentration,
and other PK statistics.
In the first chunk rats are limited parameterized as rats (for _fup_ and _CLint_), and in
the second chunk human parameterization of _fup_ and _CLint_ are allowed.
